---
title: "A *very* important report on global UFO sightings and abductions"
subtitle: "*(But mainly a lesson on using R Markdown and GitHub Pages to share your work)*"
author: "Compiled on `r date()` by Sam Csik & Juliette Verstaen"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: spacelab
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br> 

## About and Credits

These materials were developed by Sam Csik & Juliette Verstaen using UFO Sighting Data from the National UFO Reporting Center (NUFORC). Data can be downloaded from [kaggle](https://www.kaggle.com/NUFORC/ufo-sightings). Source code can be found on [GitHub](https://github.com/samanthacsik/RLadiesSB-github-pages).

In this example report, we demonstrate how to use a variety of different `HTML widgets` that allow for interactive web visualizations, as well as how to leverage [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [GitHub Pages](https://pages.github.com/) for building web pages to display reports (or whatever else you might dream of creating with RMarkdown).

Expand hidden `Code` chunks to see how we've created the following tables, maps, and plots. Also, be on the lookout for [**TIP**]{style="color: green;"}s and [**RESOURCE**]{style="color: blue;"}s throughout for pointers and links to resources!

And finally, before we get started, the [R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/) by *Yihui Xie, Christophe Dervieux, & Emily Riederer* and [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/) by *Yihui Xie, J. J. Allaire, & Garrett Grolemund* contain **tons** of incredible information and answers to nearly all of the R Markdown-related questions we have encountered. We recommend coming back to these to learn more about all-things R Markdown.

<br> 

## Setting up your RMarkdown file

-   [customizing the YAML](linked_pages/yaml.html)
-   [a note on code chunks](linked_pages/code_chunks.html)
-   [markdown syntax](linked_pages/markdown_syntax.html)
-   [creating tabs](linked_pages/creating_tabs.html)
-   [publishing your .Rmd as a GitHub Page](linked_pages/github_pages.html)

```{r, warning = FALSE, message = FALSE, results = 'hide'}
###########################
# load packages
###########################

# for cleaning and wrangling
library(tidyverse)
library(janitor)
library(lubridate) 

# for map making and general spatial work
library(leaflet)
library(mapview)
library(sf)

# for working with time series data
library(dygraphs)
library(xts)

# for making tables
library(DT)
library(gt)

# for palette making
library(RColorBrewer)
library(colorspace)
library(scales)

# for reproducibility
library(here)

##############################
# source custom alien color palette
##############################

source(here::here("code", "palettes.R"))

###########################
# load data
###########################

UFO_data_all <- read_csv(here::here("data", "UFO_complete.csv"))

###########################
# bunch of parsing failures from columns that are shifted; we're going to just filter those out bc they don't have appropriate lat/long data
###########################

# see parsing failures here
UFO_parsing_failures <- UFO_data_all %>% 
  filter(is.na(city))

# remove those from the dataset
UFO_data_cleaned <- UFO_data_all %>% 
 filter(!is.na(city))
```

<br>

## Alien Encounters {.tabset}

### Timeseries (`dygraph`)
```{r,  warning = FALSE, message = FALSE}
##############################
# get colors from our custom alien palette
##############################

colors <- alien_palette("galaxy_quest", 8, "discrete")

##############################
# wrangle data
##############################

UFO_dygraph <- UFO_data_cleaned %>%
  clean_names() %>%
  select(-date_posted) %>%
  separate(datetime, c("month", "second", "third"), sep = "/") %>%
  separate(third, c("year", "time"), sep = " ") %>%
  group_by(year) %>%
  count() %>%
  rename(number_instances = n) %>%
  filter(!is.na(year)) %>%
  mutate(date_seen = paste("01/01/", year, sep = "")) %>%
  mutate(date_seen = as.character(date_seen),
         date_seen = as.Date(date_seen, format = "%m/%d/%Y"))

##############################
# create time series object
##############################

ufo_timeseries <- xts(x = UFO_dygraph$number_instances,
                              order.by = UFO_dygraph$date_seen)

##############################
# finally create the plot
##############################

dy_plot <- dygraph(ufo_timeseries, main = "Aliens from Earth")%>%
  dyRangeSelector() %>%
  dyOptions(labelsUTC = TRUE, fillGraph=TRUE, fillAlpha=0.1, drawGrid = FALSE, colors= colors[1]) %>%
  dyCrosshair(direction = "vertical") %>%
  dyHighlight(highlightCircleSize = 5, highlightSeriesBackgroundAlpha = 0.2, hideOnMouseOut = FALSE) %>%
  dyRoller(rollPeriod = 1) %>%
  dyAxis("y", label = "Number of sightings, abductions, etc")

##############################
# display plot in knitted doc
##############################

dy_plot

```

[**RESOURCE:**]{style="color: blue;"} For more information and tips on creating dygraphs, check out [dygraphs for R](https://www.htmlwidgets.org/showcase_leaflet.html).

<br> 

### Mapping duration (`mapview`)

```{r,  warning = FALSE, message = FALSE}
##############################
# clean data
##############################

UFO_time <- UFO_data_cleaned %>%
  clean_names() %>%
  select(-date_posted) %>%
  mutate(longitude = as.numeric(longitude),
         latitude = as.numeric(latitude)) %>%
  mutate(duration_length = case_when(duration_seconds <= 60 ~ "short",
                                     duration_seconds >= 360 ~ "long",
                                     T ~ "medium")) %>%
  filter(!is.na(longitude),
         !is.na(latitude)) %>%
  group_by(duration_length, longitude, latitude) %>%
  count() %>%
  rename(number_instances = n)

##############################
# need to convert to a spatial feature object
##############################

UFO_spatial <-  st_as_sf(UFO_time, coords = c("longitude", "latitude"),
           crs = 4326)

##############################
## create the map!
##############################

UFO_map_mv <- mapview(UFO_spatial, zcol = "duration_length", cex = "number_instances", alpha = 0.2)

##############################
# display map in knitted doc
##############################

UFO_map_mv
```

[**RESOURCE:**]{style="color: blue;"} Explore Jamie Montgomery's [mapview and mapedit tutorial from SB Eco-Data Science](https://github.com/jamiecmontgomery/mapview-mapedit) for more mapview fun.

<br>

### Spaceship shapes (`gt`)

```{r,  warning = FALSE, message = FALSE}
##############################
# get colors from our custom alien palette
##############################

colors_table <- alien_palette("galaxy_quest", 8, "discrete")

##############################
# wrangle data (get counts by UFO shape)
##############################

UFO_shape_count <- UFO_data_cleaned %>%
  clean_names() %>%
  select(-date_posted) %>%
  mutate(longitude = as.numeric(longitude),
         latitude = as.numeric(latitude)) %>%
  group_by(shape) %>%
  count() %>%
  rename(number_instances = n) %>%
  filter(!is.na(shape)) ## There are 28 unique shapes

##############################
# wrangle data (get total time each shape was viewed)
##############################

UFO_shape_times <- UFO_data_cleaned %>%
  clean_names() %>%
  select(-date_posted) %>%
  mutate(longitude = as.numeric(longitude),
         latitude = as.numeric(latitude)) %>%
  group_by(shape) %>%
  dplyr::summarise(total_seconds = sum(duration_seconds, na.rm = TRUE)) %>%
  filter(!is.na(shape))

##############################
# combine data frames
##############################

UFO_shape <- left_join(UFO_shape_count, UFO_shape_times) %>%
  mutate(shape_rate_seconds = total_seconds/number_instances) %>%
  mutate(Hours = hour(seconds_to_period(shape_rate_seconds)),
         Minutes = minute(seconds_to_period(shape_rate_seconds)),
         Seconds = round(second(seconds_to_period(shape_rate_seconds)), digits = 0)) %>%
  unite(shape_rate, c("Hours", "Minutes", "Seconds"), sep = ":") %>%
  ungroup()

##############################
# make the `gt` table
##############################

Table1_UFO_shapes <- UFO_shape %>%
  dplyr::arrange(desc(shape_rate_seconds)) %>%
  select(-shape_rate_seconds) %>%
  gt() %>%
  tab_header(title = "Alien Spaceship Shapes & Time in the Air") %>%
  cols_label(shape = "Shape",
             number_instances = "Number Sightings",
             total_seconds = "Total Seconds Observed",
             shape_rate = "Time per each Sighting")  %>%
  cols_align("center") %>%
  tab_options(
    table.background.color = colors[1]) %>%
  tab_style(
    style = list(
      cell_fill(color = colors[6]),
      cell_text(weight = "bold", color = colors[1])
    ),
    locations = cells_title(groups = "title")
  )

##############################
# display table in knitted doc
##############################

Table1_UFO_shapes
```

[**RESOURCE:**]{style="color: blue;"} Explore Juliette Verstaen's [`gt` R-Ladies workshop](https://github.com/j-verstaen/gt_workshop_SB_R_Ladies) for more table fun.

<br>

## Alien Abductions {.tabset}

### All accounts of abductions (`DT`)

```{r, warning = FALSE, message = FALSE}
##############################
# clean up UFO_data to display in an interactive table
##############################

Table2_abduction_data <- UFO_data_cleaned %>% 
  clean_names() %>% # clean up those col headers using the janitor package
  filter(str_detect(comments, "(?i)abduction")) %>%  # filter for any comments that mention 'abduction'
  select(datetime, date_posted, country, state, city, duration_seconds, shape, comments) 

##############################
# generate an interactive table (using the `DT` package) 
##############################

Table2 <- datatable(Table2_abduction_data, # your data frame
                    class = 'cell-border stripe', 
                    colnames = c("Date/Time (of encounter)", "Date Reported", # rename column names
                                 "Country", "State", "City", "Duration of Encounter (sec)", 
                                 "Shape of UFO", "Comments"),
                    caption = htmltools::tags$caption( # add a caption
                      style = 'caption-side: top; text-align: left;', # set location/alignment of caption text
                      htmltools::em('Table 1: Global accounts of alien abductions from 1910-2013.')), # write caption
                    filter = 'top', 
                    options = list(pageLength = 5, autoWidth = TRUE)) # set number of viewable rows

##############################
# display table in knitted doc
##############################

Table2
```

[**RESOURCE:**]{style="color: blue;"} Explore [DT: An R interface to the DataTables library](https://rstudio.github.io/DT/) for more ways to customize your DT tables.

<br>

### Map of alien abductions (`leaflet`)

```{r}
##############################
# wrangle data (get latitude, longitude, comment information for all comments mentioning "abduction")
##############################

leaflet_abduction_data <- UFO_data_cleaned %>% 
  filter(str_detect(comments, "(?i)abduction")) %>% # filter for any comments mentioning 'abduction'
  mutate(latitude = as.numeric(latitude), # leaflet requires lat/long to be numeric values
         longitude = as.numeric(longitude))

###########################
# make a custom icon 
###########################

UFO_icon <- makeIcon(iconUrl = "media/ufo.png", # file path to saved .png (or URL)
                     iconWidth = 35, iconHeight = 45) # adjust size

###########################
# make interactive map (using the `leaflet` package)
###########################

abduction_map <- leaflet(leaflet_abduction_data) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% # add basemap and choose style
  addMarkers(lng = ~longitude, lat = ~latitude, icon = UFO_icon, popup = ~ comments) %>% # add UFO_icon as data points and show comments in the popup
  addMiniMap(tiles = providers$CartoDB.Positron, # add a mini map to show zoom coverage
             toggleDisplay = TRUE) # allows user to collapse mini map

###########################
# display map in knitted doc
###########################

abduction_map
```

[**RESOURCE:**]{style="color: blue;"} Explore [Leaflet for R](https://rstudio.github.io/leaflet/) for more ways to customize your Leaflet maps.  

<br>


## Additional Resources

-   The [R Markdown Reference Guide](https://rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf) is a great cheat sheet for Markdown syntax and knitr chunk options
-   [R Markdown Cheat Sheet](https://rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf) (also find more cheatsheets for all things R/RStudio by clicking Help > Cheatsheets)
-   For a list of awesome interactive **`htmlwidgets`** for R see [here](https://www.htmlwidgets.org/showcase_leaflet.html). Find even more options in the [htmlwidgets gallery](http://gallery.htmlwidgets.org/).
-   [A Guide to GitHub Pages](https://www.thinkful.com/learn/a-guide-to-using-github-pages/). This is a cool choose-your-own-adventure-style resource if you're comfortable in the command line.
- For creating your own palette, I use the ColorPick Eyedropper extension (for [Chrome](https://chrome.google.com/webstore/detail/colorpick-eyedropper/ohcpnigalekghcmgcdcenkpelffpdolg?hl=en), for [Firefox](https://addons.mozilla.org/en-US/firefox/addon/colorpick-eyedropper/), but any color picking tool will work)
